---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clccommunities

<!-- badges: start -->
<!-- badges: end -->

The goal of `clccommunities` is to make it easy to identify postal codes in low income or rural communities. 

## Installation

You can install the latest version of the `clccommunities` package from [GitHub](https://github.com). 

``` r
# Run this line once to install the devtools package
# install.packages("devtools") 

# Run this line once to install or update the clccommunities package
devtools::install_github("context-dependent/clccommunities")
```

## Define Low Income and Rural Communities in Existing Data


```{r example-existing}
library(tidyverse)
library(clccommunities)
## basic example code

ontario_schools_plus <- ontario_schools %>% 
  
  mutate(
    clc_low_income = clc_calculate_low_income(postal_code), 
    clc_rural = clc_calculate_rural(postal_code)
  )

ontario_schools_plus %>% 
  count(clc_low_income) %>% 
  knitr::kable()

ontario_schools_plus %>% 
  count(clc_rural) %>% 
  knitr::kable()
```

## Add All Columns from the Reference Dataset to Existing Data

```{r example-reference}
# Run this once to install the tidyverse family of packages.
# install.packages("tidyverse")

library(tidyverse)
library(clccommunities)

ontario_schools_community_data <- add_community_data(ontario_schools, postal_code)

# Run this once to install the skimr package (not necessary, but helpful in general)
# install.packages("skimr")

skimr::skim_without_charts(ontario_schools_community_data)
```

## Making a Simple Choropleth 

FSA polygons are available in the `fsa_shp` package data.
The initial installation process for the `sf` 
([Simple Features](https://r-spatial.github.io/sf/)) package is 
slightly more involved than the process for the other packages used in these examples.
You will need to first install the geospatial libraries that `sf` interfaces with on your computer. 

```{r example-choropleth}

library(sf)
library(tidyverse)
library(clccommunities)

schools_per_fsa <- ontario_schools %>% 
  
  add_community_data(postal_code) %>% 
  group_by(fsa) %>% 
  summarize(
    n_schools = n()
  )
  
schools_per_fsa_shp <- fsa_shp %>% 
  
  left_join(fsa_data, by = "fsa") %>% 
  filter(province == "Ontario") %>% 
  left_join(schools_per_fsa, by = "fsa") %>% 
  mutate(
    n_schools = coalesce(n_schools, 0)
  )

n_schools_range <- range(schools_per_fsa_shp$n_schools)

schools_per_fsa_shp %>% 
  
  ggplot() + 
  geom_sf(aes(fill = n_schools), colour = NA) + 
  scale_fill_viridis_b(
    guide = guide_colorbar(
      title = "# of Schools", 
      barheight = 12, 
      barwidth = 0.5
    ), 
    breaks = seq(0, 65, by = 10)
  ) + 
  theme_void() 
  
```

## Using package data outside of R

To use these data outside of the R environment, write the `fsa_data` object to a 
csv file, which you can open in Excel, Google Sheets, or any other program. 
```{r export}
library(readr)
library(clccommunities)

write_csv(fsa_data, "path/to/destination-folder/file-name-of-your-choice.csv")
```

## Data Sources

[June 2021 Ontario Public School Address File](https://data.ontario.ca/dataset/ontario-public-school-contact-information) 

[2016 Census Profiles](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/download-telecharger/comp/page_dl-tc.cfm?Lang=E)

[2016 Census Boundary Files](https://www12.statcan.gc.ca/census-recensement/2011/geo/bound-limit/bound-limit-2016-eng.cfm)

[2016 Geographic Attribute File](https://www12.statcan.gc.ca/census-recensement/2011/geo/ref/att-eng.cfm)


